---
title: "Yolo To Biigle"
author: "Nils"
date: "02/12/2021"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
editor_options: 
  markdown: 
    wrap: 72
---

## Before you start

This step takes place after you ran predictions with a yolov5 weight
files and you produced the 'labels' folder with predictions in .txt
files with

this means you must use these flags when calling yolov5/detect.py :
--save-txt --save-conf

\- It is better to have access to the .yaml file of the weights you use
to make your predictions but not essential

\- you can manually write the corresponding labels in the script if
easier but beware, this is harder to manage if you have a lot of labels

\- You must have URLs for your images and know how to upload them onto
biigle

\- You need the label tree raw file so you can link your verbose names
to the label IDs biigle uses.

\- This tree must be linked to the project you will modify

\- Obvious but you need internet connection to interact with the online
Biigle session

## Setting up you environment

First, load your packages, set the WD and set the pathways to the
different folders

you can run this code to create the structure in the folder. once it is
done, place your labels/annotations into the labels folder and add the
taxonomy catalogue with the Biigle IDs into the taxonomy folder

```{r include=FALSE}
library(jsonlite)
library(httr)
library(imager)
library(magick)
library(tidyverse)
library(magrittr)


# set a pathway to your WD - just so you know where it is
# the rest of the pathway can be nested here, up to you
wd <- "C:/Your/path/yoloannotations"

# make the directory where your rmd file is as your wd
  # wd  <- dirname(rstudioapi::getSourceEditorContext()$path) 

# set pathway to your predictions directory 
predictions_dir <- "C:/Repos/yolov5/runs/detect/exp"

# Optional :: pathway to folder where your images are 
images_dir <- "Z:/images/Test_images/imageset"

# set the path to where your label catalogues are
taxonomy_dir  <- "C:/your/path/taxonomy trees"

# pathway to your original data directory Where the .yaml file is
data_dir <- "C:/your/path/projName"

# set the images width and height of the images - right click on image to find out their resolution 
# !! dont get this wrong !!
    images.width <- 2448
    images.height <- 2048
# you can also use the API or Go on Biigle and point your mouse on the bottom right corner of the image to get their exact resolution
# Overall, it is the simplest way to do this and making sure you have the correct resolution

```

## set your table of Biigle names

Yolo and Biilge will use different names for each OTU. You need to
translate from one to the other There are many way to get that step. the
easiest one is to write the names

If you don't have access to the yaml file or can't be bothered to paste
the pathways here (fair...), just put down the class names as a vector
here:

```{r}

# path to the hyper parameter yaml of your Yolo model that you made predictions with
diryaml <- paste0(data_dir,"/dataset.yaml")
 

# print the names 
diryaml %>% read.delim()%>% unlist() %>% as.vector() %>% str_subset(pattern = "names:") -> v
print(v)
  v %>% str_remove(pattern =  ".*\\['|\\]" ) %>% str_remove(pattern = "\\'].*|\\]") -> v
    v %>%  str_split(pattern = "','") %>%  unlist() -> objnames
print(objnames)



# manually enter the object names list
# read it from the yaml file, it should be a line that start with "names : ['names1','names2']
# and paste it there 
# objnames <- c('names1','names2','names3','names4','names5','names6')

OBJ_NAME <- 
  tibble( yolo_name = objnames  , 
  class_id = c(1:length(objnames)) -1 )

# example of label names from change names to DSCU labels
OBJ_NAME   %<>%
  mutate(
    # replace this vector with your names
    label_name = c(
      'OTU585',
      'OTU2',
      'OTU992',
      'OTU339',
      'OTU404',
      'OTU555',
      'OTU582',
      'OTU963'
    )
  )


# change names to smartr ID labels 
# these names must be in the same order as the corresponding yolonames
OBJ_NAME   %<>% 
   # replace this vector with your names
  mutate(label_name = c('SM670 Acanella arbuscula',
                        'SM362 Ceriantharia msp1',
                        'SM1099 Craniella longipilis',
                        'SM266 Munida msp1',
                        'SM240 Eumalacostraca msp2',
                        'SM397 Phormosoma placenta',
                        'SM38 Actiniaria msp20',
                        'SM783 Bathyplotes natans') )


# you can always check and modify manually with: 
# OBJ_NAME %>% edit()



# view final names
OBJ_NAME %>% pull(label_name) %>%  print()


```

## load the YoloV5 annotations

V5 makes predictions as txt files. You need to aggregate them all into
one

```{r message=FALSE}

paste0( predictions_dir,"/labels" ) -> annotations_dir

# list the annotations files
annotations_dir %>% 
  list.files(pattern = ".txt")   -> txtfiles
# list the images - if you have them in a folder
images_dir %>% 
  list.files(pattern = ".jpg") -> imgs
# list the images - as the ones for which you have annotations (omiting empty images)
txtfiles %>%  str_replace(pattern = ".txt",replacement = ".jpg") -> imgs



Dl <- list()
length(txtfiles) -> l

for(i in seq(txtfiles) ){
  
  # open the label txt file
  txtfiles[i] -> label.i
  print(label.i)
  paste0(annotations_dir,"/", label.i) %>% read_table(col_names = F) -> d.i
  
  # change column names and add image
  d.i %<>% set_names(c("class_id", "center.x" , "center.y" , "width", "height","confidence")) %>%
    mutate(filename = label.i %>%  str_replace(pattern = "txt","jpg")) %>% 
    relocate(filename, .before = class_id)
  
  d.i -> Dl[[i]]
  
    paste0( round(i/l,digits = 2)*100," % done" ) %>%  print()
  
}

 
Dl %>%  bind_rows() -> annotations

# add a label name
annotations %<>% left_join( OBJ_NAME, by = "class_id") %>% relocate(label_name, .before = "class_id")


```

the yolo annotations are not compatible with Biigle so they have to be
converted back to Biigle's antive format where coordinates are in pixel
numbers

```{r}


annotations %<>%  
  # add the image dimensions 
  #!!!! make it a step in the upload to biigle script so you can read the image attribute there !!!!
  mutate(attrs.width = images.width, attrs.height = images.height ) %>% 
  # change names of relative coord. to make rooms for calculate pixel coodinates
    rename(
   relative_coordinates.center_x = center.x , 
   relative_coordinates.center_y = center.y,
   relative_coordinates.width = width,
   relative_coordinates.height = height
    )

# denormalise the coordinates
    annotations %<>% 
      # resize the center X dimension
      mutate(center.x = relative_coordinates.center_x *  attrs.width ) %>%
      # resize the center Y
      mutate(center.y = relative_coordinates.center_y *  attrs.height ) %>% 
      # resize the X width 
      mutate(width = relative_coordinates.width *  attrs.width ) %>% 
      # resize teh y dimension 
      mutate(height = relative_coordinates.height  * attrs.height ) 
    
# add more dimensions columns for easie coordinate calculations for other shapes
    annotations %<>% 
    mutate(xleft = center.x - (width/2), 
           ybottom = center.y - (height/2),
           xright = center.x + (width/2),
           ytop= center.y + (height/2)
    )  


```

## make a table to link the label names in yolo with the keys in the species catalogue

Biigle has a centralised mangment system for all the labels of all the
catalogues where each label has a unique id number. this lets you input
any verbose names and ensure it will not be confused with the label of
the same name in another tree. So you must retrieve that id before you
add it into biigle

```{r}

# load the species catalog

# add the taxonomy to the lables (so annotations can be grouped into higher levels bins)
paste0(taxonomy_dir,"/smartarid - Label_tree.csv") %>% read_csv() %>% 
  select(- label_tree_id) -> tree

tree %<>% arrange(id) %>%  rename(label_name = name, label_id = id, aphia_id = source_id)

# attach the tree labels to the label names table
OBJ_NAME %<>% left_join(tree, by = "label_name")


# you can always check and modify manually with: 
# OBJ_NAME %>% edit()


```

## run some checks to make sure all labels have an equivalent in catalgue

```{r}
  # make sure the names are in the tree
OBJ_NAME %>% mutate( mismatches =  OBJ_NAME$label_name %in% tree$label_name ) %>% 
  filter(mismatches == FALSE) -> d

if(nrow(d) == 0){
  print(paste0("All yolo labels are matched in the catalogue"))
}else{
  print(paste0(nrow(d)," yolo labels are not in the catalogue"))
}


# delete the matching table
rm(d)

# assuming all went well, add the label ids to the actual annotations 
annotations %>%
  left_join(OBJ_NAME ,
            by = c("label_name", "class_id", "yolo_name")) -> annotations_ids



```

## convert the yolo annotations to biigle format

first, what biigle volume will be working on?

do you have the images? this should download them in your WD

```{r echo=TRUE}
serverURL <- "https://server.com/bucket name"
paste0(wd,"/imageDL") -> imageDLdir
dir.create(imageDLdir)

for (imageName in imgs) {
  print(imageName)
  # dont download the image if it is already there
  if(!file.exists(paste0(imageDLdir,"/",imageName))){
    print(" -- Downloading -- ")
  paste0(serverURL, "/",imageName ) %>% download.file(destfile = paste0(imageDLdir,"/",imageName), mode = 'wb')
  }else{print("already downloaded")}# checking image doesnt exist before downloading 
  
}# download images to your yolo directory

```

If you dont have one yet, make a biigle volume with the images for which
you have annotations Assuming, images with the same names (and the same
animals, idealy) are uplaoded on a server somewhere, create a new volume
and link those specific images to it - Give it a name in "2. Choose a
name or import from file" box - Pop the serverURL into the "3. Choose a
file source" box - Add the images list in the "Volume files" box

```{r echo=TRUE}
# print the names of the images that will be uploaded
print(imgs)

# get the images names
writeClipboard(
    paste0(
        unlist(paste(imgs,collapse = " , ")
        ) ,
        collapse = " , "))

```

once the volume is done, get its volume id so you can point the API
towards the right images on biigle otherwise, pick an existing volume

## biigle API parameters

To interact with Biigle, you can either use the manual interface or the
API. since you already know where your annotations are in the image and
what labels they have you can use the API. You will need to know you
Biigle credential before you are authorized to interact with your
volumes. You also need a token for your identity to be tied to your
requests and authorized. You can get your token at:
<https://biigle.de/manual/tutorials/login-and-account-settings>

```{r}

# put your credentials here: 
login <- "your email address"
token <- "your token number"


# base url
base_url <- "https://biigle.de/api/v1"

# the ID of the volume you will add your images to  
Volume <- 9999 # ID code of the volume to modify


# go check your volume
shell.exec(paste0("https://biigle.de/volumes/",Volume))

```

Once you are sure where your images will go, we need to get some info
from biigle about how it calls these images so we can give it the
instructions it understands

first, get the ids of the images in r. This is also a gentle intro to
using the API When you send a request to the server it respond with a
code that means success, fail or inbetween

Some server responses Status: 201 \# Created -- that probably means
success Status: 422 \# unprocessable entity -- I guess that is a fail

all API commands are listed here:
<https://biigle.de/doc/api/index.html#api-_>

this step can take a while if there are a lot of images to get the info
about

```{r}

# get the new images infos (ID etc)
      # build the URL from the base API address, which volu
      url <- paste0(base_url, 
                    # which volume you want info from
                    "/volumes/",Volume,
                    # and what exacly you want to know about this volums
                    "/files" )

# it should look like this : "https://biigle.de/api/v1/volumes/123456789/files"

      
# Then, send the request to get a list of image IDs for that volume
      # send the request with you credential attached
      GET(url, authenticate(login, token) ) %>% 
        content(., as = "text", encoding = "UTF-8") %>% 
        jsonlite :: fromJSON( flatten = TRUE) -> imageIDs
      
tibble(filename = imgs, biigleid = imageIDs, volume_id = Volume) -> imageinfos

      
# Add that info to the annotation tables
imageinfos %>% select(biigleid  , filename, volume_id )  %>%
  left_join(annotations_ids ,by = "filename" , keep = FALSE ) %>% 
  select(-contains("relative"), - class_id, yolo_name) -> annotations_biigle



# You can get more info about the images. This is not necessary here but potentially useful to verrify hte prameters     
# build URL
  url.i <- paste0(base_url,"/images/", imageIDs[1] )
  # send request
  GET(url.i, authenticate(login, token) ) %>% 
    content(., as = "text", encoding = "UTF-8") %>% 
    # reformat 
    jsonlite :: fromJSON( flatten = TRUE) %>%
    unlist %>% t %>% as.data.frame() -> Image_infos
  
      
 
```

last but important, add some additional parameters that the API can
interpret, the shape of the annotation is most important

```{r}

# add columns for shapes and confidence
annotations_upload <- annotations_biigle %>% 
  # remove the unnecessary columns
  # add the arguments that the API will interpret to make the labels
  # !! the shape names and and IDs are essential
  mutate( shape_id = 5, # 4 is for Circle, # 5, Rectangle, #1, Point
          shape_name = "Rectangle") # names is just there for clarity

 
```

## Uploading the annotations

You should be all set to actually start uploading the images

it is good practice to run the loop once (i = 1) to see if the label
gets in the right place. Run the commands once, it should print what
image and label it used. Go to this image on your volume, there should
now be one annotation. If it looks right, delete it manually (or it will
get duplicated) and run the loop.

You may want to run the loop without actually send the requests to chek
the URLs are correctly assembled. in which case, dont give it the green
light.

by design, you can only send 10,800 requests per hour. There are a
couple of ways around this but they are not covered here.

the rp list collects the server resonse to your request. If all goes
well, it will say "201". other codes exists for when things dont work.
This is not always helpful but can help narrowing down what the issue
is.

```{r}

# upload -----------------------------------------------------------------------
# are you sure want to upload? only give green light if you are certain the code is right 
# errors could have unexpected consequences
upload_enabled <- "Green" # "Green" or "notGreen

# make sure there are not too many annotations 
if( annotations_upload %>% nrow() > 10799){print("Too many annotations, this will not go through")}
 
# server responses to annotation input - - - - - - - - - - - - - -- - - - - -  -
upload_response <- list()

for (i in 1:nrow(annotations_upload)) {
  
  annotations_upload %>% slice(i) -> A.I
  cat( paste( A.I$filename," (id:", A.I$biigleid, ") - - - " , A.I$label_name), "in" , A.I$shape_name)
   
# build the bounding box  - that is the coordinates of the rectangles
  topleft = c(A.I$xleft, A.I$ybottom)  
  bottomleft = c(A.I$xleft, A.I$ytop)
  bottomright = c(A.I$xright, A.I$ytop) 
  topright = c(A.I$xright, A.I$ybottom)
  
biigle_coordinates = c(topleft, bottomleft, bottomright, topright)
  
  paste0(base_url,"/images/", A.I$biigleid,"/annotations" ) -> url.i
  # make the query list for each annotation
  # these are the required parameters
  list( shape_id = A.I$shape_id, # 4 is for circle
        label_id = A.I$label_id,
        # add confidence number - Either the one from yolo
        confidence = A.I$confidence,
        # or, an arbitrary number
        # confidence = 0.75 ,
        # make the list of points coordinates
        points = biigle_coordinates ) -> body_list
  
  
  # send the request (if enabled)
  if(upload_enabled == "Green"){
    cat("Sending request to server - - ")
    # say that to the API
    POST(url.i,
         config = authenticate(login, token),
         body = body_list, 
         encode = "json"    ) -> rp
    # example image
    rp -> upload_response[[length(upload_response) +1]]
  }# 
  print("Next annotation") 
  rm(url.i)
  
}      
      


```
